Automating Verification of MPC Security: $\metaprot$ to $\minifed$

### 2024-04-19 revision-1/lang.tex
checkpoint

\subsection{Example: Passive-Secure Addition}

@@

To capture this distinction we introduce a function $\otp$
with the following specification:
$$
\delta(\otp,v_1,v_2) \defeq v_1 \fplus v_2
$$
Although the semantics are the same as addition, the use of $\otp$
makes a declarative distinction, but more importantly we will be
able to assign a type to $\otp$ that enforces the one-time discipline
on its first argument via type linearity as will be discussed in Section
\ref{section-types}.

@@

$$
(v, [m_1,\ldots,m_n])
$$
$$
k_1,\ldots,k_n
$$
$$
m_\cid = k_\cid + (k_\Delta * v)
$$

$$
\delta(\macplus,(v^1, [m_1^1,\ldots,m^1_n]),(v^2, [m_1^2,\ldots,m^2_n]))
\defeq
(v^1 \fplus v^2, [m_1^1 \fplus m_1^2 ,\ldots,m^1_n \fplus m_n^2])
$$

$$
\elab{\macv{\elab{v}{\cid}}}{1} \macplus \cdots \macplus \elab{\macv{\elab{v}{\cid}}}{n} =
(v,\ldots)
$$

@@

$$
\delta(\macauth, (v, [\ldots,m_\cid,\ldots]), k_\cid) \defeq
     (v, [\ldots,m_\cid,\ldots]) \text{\ if\ } m_i = k_i + (k_\Delta * v)
$$

$$
\begin{array}{lcl}
  \elab{\mesg{a}}{2} &:=&
  \elab{(\macotp\ \macv{\elab{\secret{x}}{1}}\ \macv{\elab{\flip{a}}{\Oracle}})}{1}\\
  \elab{\mesg{a}}{1} &:=&
  \elab{(\macotp\ \macv{\elab{\secret{x}}{1}}\ \macv{\elab{\flip{a}}{\Oracle}})}{2}\\
  \elab{\mesg{b}}{2} &:=&
  \elab{(\macotp\ \macv{\elab{\secret{y}}{2}}\ \macv{\elab{\flip{b}}{\Oracle}})}{1}\\
  \elab{\mesg{b}}{1} &:=&
  \elab{(\macotp\ \macv{\elab{\secret{y}}{2}}\ \macv{\elab{\flip{b}}{\Oracle}})}{2}\\
  \elab{\mesg{d}}{1} &:=&
  \elab{(\macauth(\mesg{a}, \mack{\elab{\secret{x}}{1}}{2} \fminus \mack{\elab{\flip{a}}{\Oracle}}{2}) \macplus (\macv{\elab{\secret{x}}{1}}\macminus\macv{\elab{\flip{a}}{\Oracle}}))}{1}\\
  \elab{\mesg{e}}{1}&:=&
  \elab{(\macauth(\mesg{b}, \mack{\elab{\secret{y}}{2}}{2} \fminus \mack{\elab{\flip{b}}{\Oracle}}{2}) \macplus (\macv{\elab{\secret{y}}{2}}\macminus\macv{\elab{\flip{b}}{\Oracle}}))}{1}\\
  \rvl{1} &:=&
  \elab{( (\mesg{d} \mactimes \mesg{e}) \macplus
          (\mesg{d} \mactimes \macv{\elab{\flip{b}}{\Oracle}}) \macplus
          (\mesg{e} \mactimes \macv{\elab{\flip{a}}{\Oracle}}) \macplus \macv{\elab{\secret{c}}{\Oracle}}
    )}{1}\\
  \elab{\mesg{d}}{2} &:=&
  \elab{(\macauth(\mesg{a}, \mack{\elab{\secret{x}}{1}}{1} \fminus \mack{\elab{\flip{a}}{\Oracle}}{1}) \macplus (\macv{\elab{\secret{x}}{1}}\macminus\macv{\elab{\flip{a}}{\Oracle}}))}{2}\\
  \elab{\mesg{e}}{2}&:=&
  \elab{(\macauth(\mesg{b}, \mack{\elab{\secret{y}}{2}}{1} \fminus \mack{\elab{\flip{b}}{\Oracle}}{1}) \macplus (\macv{\elab{\secret{y}}{2}}\macminus\macv{\elab{\flip{b}}{\Oracle}}))}{2}\\
  \rvl{2} &:=&
  \elab{( (\mesg{d} \mactimes \mesg{e}) \macplus
          (\mesg{d} \mactimes \macv{\elab{\flip{b}}{\Oracle}}) \macplus
          (\mesg{e} \mactimes \macv{\elab{\flip{a}}{\Oracle}}) \macplus \macv{\elab{\secret{c}}{\Oracle}}
    )}{2}\\
  \out{1} &:=& \elab{(\rvl{1} \macplus \macauth(\rvl{2},\ldots))}{1} \\
  \out{2} &:=& \elab{(\macauth(\rvl{1},\ldots) \macplus \rvl{2})}{2}
\end{array}
$$

>>>>

\subsection{Example: Passive Secure Addition}

@@

To capture this distinction we introduce a function $\otp$
with the following specification:
$$
\delta(\otp,v_1,v_2) \defeq v_1 \fplus v_2
$$
Although the semantics are the same as addition, the use of $\otp$
makes a declarative distinction. But more importantly we will be
able to assign a type to $\otp$ that enforces the one-time discipline
on its first argument via type linearity, and reflects hiding of
its second argument via security type labels, as will be discussed in Section
\ref{section-types}.

@@

\begin{fpfig}[t]{Top-to-bottom: Extended $\minifed$ syntax with BDOZ-style shares,
    interpretation of operations, and pre-processing properties of memories.}{fig-bdoz}
  {
$$
\begin{array}{rcl@{\hspace{8mm}}r}
  \multicolumn{3}{l}{m,k \in \mathbb{Z}_p \qquad \macv ::= (v,[m_1,\ldots,m_n])} &\\[2mm]
  %\textit{MACed\ values}\\[2mm]
  \be &::=& \cdots \mid \macv \mid \be \macplus \be \mid \be \mactimes \be \mid \be \macminus \be\\[2mm]
  x &::=& \cdots \mid \macx{\secret{w}}{\cid} \mid \macx{\flip{w}}{\cid} \mid \mack{x}{\cid}
\end{array}
$$

$$
\begin{array}{rcl}
  \fcod{(v^1, [m_1^1,\ldots,m^1_n]) \macplus (v^2, [m_1^2,\ldots,m^2_n])} &=&
  (v^1 \fplus v^2, [m_1^1 \fplus m_1^2 ,\ldots,m^1_n \fplus m_n^2])\\[2mm]
  \fcod{(v^1, [m_1^1,\ldots,m^1_n]) \mactimes v} &=&
   (v^1 \ftimes v, [m_1^1 \ftimes v ,\ldots,m^1_n \ftimes v])
\end{array}
$$
$$
\begin{array}{rcl}
  \lcod{\store, \be_1 \macplus \be_2}{\cid} &=&
  \fcod{\lcod{\store,\be_1}{\cid} \macplus \lcod{\store,\be_2}{\cid}}\\[2mm]  
  \lcod{\store, \be_1 \mactimes \be_2}{\cid} &=&
  \fcod{\lcod{\store,\be_1}{\cid} \mactimes \lcod{\store,\be_2}{\cid}}
\end{array}
$$

\begin{mathpar}
  \inferrule
      {\store(x) = v \\ \store(\macx{x}{\cid_1}) = (v_1,\ldots)\  \cdots\ 
        \store(\macx{x}{\cid_n}) = (v_n,\ldots)}
      {v = \fcod{v_1 \fplus \cdots \fplus v_n}}

  \inferrule
      {\store(\macx{x}{\cid'}) = (v,[m_1,\ldots,m_n]) \\  \cid \in [1..n]}
      {m_\cid = \fcod{\store(\mack{x}{\cid}) \fplus (\store(\elab{\flip{\ttt{delta}}}{\cid'}) \ftimes v)}}

  \store(\elab{\secret{c}}{\Oracle}) = \lcod{\store,\flip{a} \ftimes \flip{b}}{\Oracle}
\end{mathpar}
}
\end{fpfig}

$$
m_\cid = k_\cid + (k_\Delta * v)
$$

@@

$$
\delta(\macauth, (v, [\ldots,m_\cid,\ldots]), k_\cid) \defeq
     (v, [\ldots,m_\cid,\ldots]) \text{\ if\ } m_\cid = k_\cid + (k_\Delta * v)
$$

$$
\begin{array}{lcl}
  \elab{\mesg{a}}{2} &:=&
  \elab{(\macotp\ \macgv{\elab{\secret{x}}{1}}\ \macgv{\elab{\flip{a}}{\Oracle}})}{1}\\
  \elab{\mesg{a}}{1} &:=&
  \elab{(\macotp\ \macgv{\elab{\secret{x}}{1}}\ \macgv{\elab{\flip{a}}{\Oracle}})}{2}\\
  \elab{\mesg{b}}{2} &:=&
  \elab{(\macotp\ \macgv{\elab{\secret{y}}{2}}\ \macgv{\elab{\flip{b}}{\Oracle}})}{1}\\
  \elab{\mesg{b}}{1} &:=&
  \elab{(\macotp\ \macgv{\elab{\secret{y}}{2}}\ \macgv{\elab{\flip{b}}{\Oracle}})}{2}\\
  \elab{\mesg{d}}{1} &:=&
  \elab{(\macauth(\mesg{a}, \mack{\macx{\secret{x}}{1}}{2} \fminus \mack{\macx{\flip{a}}{\Oracle}}{2}) \macplus (\macgv{\elab{\secret{x}}{1}}\macminus\macgv{\elab{\flip{a}}{\Oracle}}))}{1}\\
  \elab{\mesg{e}}{1}&:=&
  \elab{(\macauth(\mesg{b}, \mack{\macx{\secret{y}}{2}}{2} \fminus \mack{\macx{\flip{b}}{\Oracle}}{2}) \macplus (\macgv{\elab{\secret{y}}{2}}\macminus\macgv{\elab{\flip{b}}{\Oracle}}))}{1}\\
  \rvl{1} &:=&
  \elab{( (\mesg{d} \mactimes \mesg{e}) \macplus
          (\mesg{d} \mactimes \macgv{\elab{\flip{b}}{\Oracle}}) \macplus
          (\mesg{e} \mactimes \macgv{\elab{\flip{a}}{\Oracle}}) \macplus \macgv{\elab{\secret{c}}{\Oracle}}
    )}{1}\\
  \elab{\mesg{d}}{2} &:=&
  \elab{(\macauth(\mesg{a}, \mack{\macx{\secret{x}}{1}}{1} \fminus \mack{\macx{\flip{a}}{\Oracle}}{1}) \macplus (\macgv{\elab{\secret{x}}{1}}\macminus\macgv{\elab{\flip{a}}{\Oracle}}))}{2}\\
  \elab{\mesg{e}}{2}&:=&
  \elab{(\macauth(\mesg{b}, \mack{\macx{\secret{y}}{2}}{1} \fminus \mack{\macx{\flip{b}}{\Oracle}}{1}) \macplus (\macgv{\elab{\secret{y}}{2}}\macminus\macgv{\elab{\flip{b}}{\Oracle}}))}{2}\\
  \rvl{2} &:=&
  \elab{( (\mesg{d} \mactimes \mesg{e}) \macplus
          (\mesg{d} \mactimes \macgv{\elab{\flip{b}}{\Oracle}}) \macplus
          (\mesg{e} \mactimes \macgv{\elab{\flip{a}}{\Oracle}}) \macplus \macgv{\elab{\secret{c}}{\Oracle}}
    )}{2}\\
  \out{1} &:=& \elab{(\rvl{1} \macplus \macauth(\rvl{2},\ldots))}{1} \\
  \out{2} &:=& \elab{(\macauth(\rvl{1},\ldots) \macplus \rvl{2})}{2}
\end{array}
$$
